---
title: "STAT340_project"
date: "12/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(maps)
library(mapdata)
library(usdata)
```

## R Markdown

```{r load the data}
data <- read.csv("Levels_Fyi_Salary_Data.csv")
head(data)
```


```{r}
# filter gender, drop NA and other in gender
filted_data <- data %>%
  filter(gender %in% c("Female", "Male"))
filted_data %>%
        group_by(gender) %>%
        summarise(n= n())


```
```{r}
# modify location column, drop location outside US and leave only state information
filted_data = filted_data %>%
  select(company, level, title, totalyearlycompensation, location, yearsofexperience, yearsatcompany, basesalary,
         stockgrantvalue, bonus, gender, Race, Education) %>% 
  group_by(location) %>% 
  mutate(new_location = strsplit(location, ', ')[[1]][2])



states <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DC", "DE", "FL", "GA", 
          "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", 
          "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", 
          "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 
          "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY")

filted_data <- filted_data %>% 
  filter(new_location %in% states)

head(filted_data)

```

```{r}
filted_data1 <- filted_data %>% 
  filter(!is.na(Race))
ggplot(filted_data1, aes(x = Race, y = basesalary, fill = gender)) +
  geom_boxplot()+ggtitle("Relationship between gender, race and basesalary")+ylab("Basesalary")
```

```{r}
ggplot(filted_data) +
  geom_boxplot(aes(x=as.character(new_location),y=basesalary, fill=new_location),
                outlier.shape = NA,alpha= 0.3) +
  theme(legend.position="none") +
  ylim(0, 260000) +
  ylab("base salary") +
  xlab("States") +
  scale_x_discrete(guide = guide_axis(angle = 60)) +
  ggtitle("Base Salary in Different State")
```

```{r}
state_mean=filted_data %>%
  group_by(new_location) %>%
  summarise(n= n(),
            mean_salary_state=mean(basesalary))
 
ggplot(state_mean, aes(x=reorder(new_location, mean_salary_state),y=mean_salary_state)) +
  geom_point() + 
  geom_line() +
  scale_x_discrete(guide = guide_axis(angle = 60)) +
  ylab("Average Base Salary") +
  xlab("States") +
  ggtitle("Average Base Salary of each States")
```

```{r}

library(lubridate)
states <- map_data("state") %>% 
  select(-subregion) %>% 
  mutate(new_location = state2abbr(region))

states = full_join(states, state_mean, by = "new_location") 

ggplot(states, aes(x = long, y = lat, group = group, fill = mean_salary_state)) +
  geom_polygon(color = "white") +
  coord_fixed(1.5) +
  theme_bw()+ ggtitle("Average Base Salary for different US States") + 
  xlab("Longitude")+ 
  ylab("Latitude")+
  labs(caption = "Tips: No data for SD(South Dakota)")+
  scale_fill_continuous(name="Average Base Salary", 
            low = "lightblue", high = "darkblue",limits = c(60000,180000), 
            breaks=c(60000,80000,100000,120000,140000,160000), na.value = "grey50")

```
```{r}
filted_data %>% 
  ggplot(aes(x = gender, y = log(basesalary), fill = gender)) +
  geom_boxplot()+ggtitle("Relationship between gender and basesalary")+ylab("Basesalary")
```
```{r}
ggplot(filted_data) + 
  geom_density(aes(x = log(basesalary), color = gender), alpha = 0.2) +
  xlab('Base Salary') +
  ylab('Density') +
  ggtitle('Density of Base Salary')
```
```{r}
# using wilcoxon test to test whether the median income of the male and female
res <- wilcox.test(basesalary ~ gender, filted_data,var.equal = TRUE)
res
```
```{r}
# We'll adapt the permutation testing approach we saw in lecture.
# Each MC replicate, we will randomly assign the data to two groups and compare their me ans.
# Then we'll see how the true difference in means compares to these.
male_data <- filted_data %>% 
  filter(gender == "Male")
female_data <- filted_data %>% 
  filter(gender == "Female")
Nrep <- 1000; 
reps <- rep( NA, Nrep)
n <- nrow(filted_data);

n_male <- nrow(male_data)
n_female <- nrow(female_data)

for( i in 1:Nrep ) {
  inds <- sample(1:n, n_male, replace=FALSE );
  reps[i] <- median(filted_data[inds,]$basesalary) - mean(filted_data[-inds,]$basesalary)
  }
# Form the CI
CI <- as.vector( quantile(reps, probs = c(0.025,0.975)) );
# Note that we could get these means with a more graceful GroupBy operation,
# But I want to make it very clear what we are computing here. 
Female_inds <- (filted_data$gender== "Female");
Female_salary <- filted_data[Female_inds,]$basesalary;
Female_median <- median(Female_salary);
Male_inds <- (filted_data$gender== "Male");
Male_salary <- filted_data[Male_inds,]$basesalary;
Male_median <- median(Male_salary);
popdiff <- Female_median-Male_median;

(CI[1] < popdiff) & (popdiff < CI[2]);
```
```{r}
mlm <- lm(basesalary ~ 1+yearsofexperience+yearsatcompany+title+new_location,male_data)
summary(mlm)
flm <- lm(basesalary ~ 1+yearsofexperience+yearsatcompany+title+new_location,female_data)
summary(flm)
```

```{r}
filted_data$company = toupper(filted_data$company)
companies <- filted_data %>% group_by(company)%>%
  summarise(n= n())

com_rela <- filted_data %>%
  group_by(company,gender)%>%
  summarise(basesalary=median(basesalary)) %>%
  pivot_wider(names_from = gender, values_from = basesalary) %>%
  drop_na() %>%
  mutate(sal_diff= Male -Female)
hist(log(com_rela$sal_diff))
mean(com_rela$sal_diff)
```

Is there any difference in gender gaps for different job positions?
```{r gender gap in different job titles, message=FALSE}
title_diff <- filted_data %>% 
  group_by(title, gender) %>% 
  summarise(basesalary=median(basesalary))

basesalary <- title_diff$basesalary
tmp <- basesalary[2:30]
tmp[30] <- 0
title_diff$basesalary <- title_diff$basesalary - tmp
gap <- title_diff %>% 
  filter(gender == "Female") %>% 
  select(-2) %>% 
  mutate(basesalary = basesalary * (-1))

ggplot(gap, aes(x=title, y=basesalary)) +
  geom_col() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  ylab("Median Base Salary Difference") +
  xlab("Job Titles") +
  ggtitle("Median Base Salary Difference in Different Job Titles (M-F)")

ggplot(gap, aes(x=basesalary)) +
  geom_histogram(color="white", fill="steelblue", bins=6) +
  xlab("Median Base Salary Difference (Male - Female)") +
  ylab("Count") +
  ggtitle("Median Base Salary Difference in Different Job Titles")
```

***

From this histogram, we can see that for most job titles, male workers have a higher median base salary than female workers. For some job titles like Data Scientist, Recruiter, the difference is very small. For Management Consultant, Human resources, and Sales, the difference is relatively large. Therefore, it seems that, within each job title, there exists base salary differences of males and females.

***

***
To verity that base salary differences of males and females in each job title is not the same, we conduct a chi-squared test to verity it. If there is no obvious gender base salary gap in each job level, the standard deviation of the base salary difference should be very small. And we consider the a standard deviation less than 500 to be small. 

Our null hypothesis is 
$H_0: $ No obvious gender base salary gap difference between each job title 
and $H_1:$ There exists gender base salary gap difference between each job title 
***

```{r echo=FALSE, message = FALSE}
library(EnvStats)
varTest(x = gap$basesalary, sigma.squared =  1/10000)
```

***
According to the chi-squared test, because the p-value is smaller than `2.2e-16`, which means the we can reject the null hypothesis that there is no obvious gender base salary gap difference between each job title.

Then we want to figure out how big the gender gap in base salary for different job titles. So for each job title, we conduct a two sample wilcoxon test to test whether the median base salary of male and female with that job title is equal. Then we show the test result as follows:
***

```{r}
job_title <- gap$title
results <- data.frame('Row'=rep(1:length(job_title)),
                      'Job_Title' = rep(NA, length(job_title)),
                      'P_Value' = rep(NA, length(job_title)))

for (i in 1:length(job_title)) {
  tmp <- filted_data %>% 
    filter(title == job_title[i])
  res <- wilcox.test(basesalary ~ gender, tmp,var.equal = TRUE)
  results[i,]$P_Value = res$p.value
  results[i,]$Job_Title = job_title[i]
}
results <- results %>% 
  mutate(significant = (P_Value <= 0.01))

results
```
***
According to the results of our wilcoxon test, we find that we can reject the null hypothesis (Male workers have the same median base salary as female workers with the same job title) for 5 job titles. For the rest 10 job titles, we cannot reject the null hypothesis. 
***

```{r echo=FALSE}
results %>% 
  filter(significant == TRUE)
```

***
In conclusion, gender gap presents differently in different job titles and in some jobs, there might not be a gender gap. From the above histogram and wilcoxon test, we find that there is strong evidence that gender gap exists in employees with job title "Product Designer", "Product Manager", "Software Engineer", "Solution Architect", and "Technical Program Manager".  

***